---
title: "p8105_hw2_wc2750"
author: "Weixi Chen"
date: "9/27/2020"
output: github_document
---

```{r message = FALSE}
# load library
library(tidyverse)
library(readxl)
library(dplyr)
```

## Problem 1

First, define a path to the dataset
```{r}
path_to_data = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx"
```

Read the Mr. Trashwheel dataset
```{r}
trashwheel_df = 
  read_xlsx(path = path_to_data,
            sheet = "Mr. Trash Wheel",
            range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data for 2018 and 2017
```{r}
precip_2018 = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet = "2018 Precipitation",
             skip = 1
            ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate (year = 2018) %>%
  relocate (year)

precip_2017 = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet = "2017 Precipitation",
             skip = 1
            ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate (year = 2017) %>%
  relocate (year)
```

Now combine annual precipitation
```{r}
month_df =
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017)

left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. The total precipitaion in 2018 is `r sum(precip_2018$total)`. The median number of sports balls in a dumpster in 2017 is `r median(trashwheel_df$sports_balls[trashwheel_df$year == 2017])`.

## Problem 2

Read and clean the NYC Transit dataset
```{r message = FALSE}
nyctransit_df = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, route1:route11, entrance_type, entry, vending, ada) %>%
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))

# preview nyctransit_df
nyctransit_df
```

The NYC Transit dataset includes variables: `r colnames(nyctransit_df)`. As so far, I've cleaned the columns' names, selected the wanted columns, and converted the entry variable from character (YES vs NO) to a logical variable. The dimension of this dataset is `r nrow(nyctransit_df)` rows * `r ncol(nyctransit_df)` columns.

Additional information about NYC Transit dataset
```{r}
# reformat distinct stations dataset
diststation_df = distinct(nyctransit_df, line, station_name, .keep_all = TRUE)
# preview diststation_df
diststation_df
# number of distinct stations
nrow(diststation_df)
# number of stations are ADA compliant
nrow(filter(diststation_df, ada == "TRUE"))
# proportion of station entrances/exits without vending allow entrance
sum(nyctransit_df$vending == "NO" & nyctransit_df$entry == "TRUE")/sum(nyctransit_df$vending == "NO")
```

Reformat data with distinct route number and route name
```{r}
# reformat distinct route number and route name dataset for distinct stations
route_diststation_df = diststation_df %>%
  # convert route8:route11 to character variable
  mutate_at(vars(route8:route11), as.character) %>% 
  pivot_longer(
  route1:route11,
  names_to = "route",
  values_to = "train")
# preview route_diststation_df
route_diststation_df
```

Additional information about serving A train
```{r}
# number of distinct stations serve the A train
nrow(filter(route_diststation_df, train == "A"))
# number of distinct stations serve the A trian and are ADA compliant
nrow(filter(route_diststation_df, train == "A" & ada == "TRUE"))
```

## Problem 3

Read and clean pols-month dataset
```{r message = FALSE}
pols_month_df =
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  # break up the variable mon into integer variable year, month, and day
  separate(mon, c("year", "month", "day"), convert = TRUE) %>%
  # replace month number with month name
  mutate (month = month.name[month]) %>%
  # create a president varaible taking values gop and dem
  mutate (president = ifelse(prez_gop == 1, "gop", "dem")) %>%
  # remove prez_dem, prez_gop, and day variable
  select(-prez_dem, -prez_gop, -day) %>%
  # organize year, month, and president as the leading columns
  relocate(year, month, president)

# preview pols_month_df
pols_month_df
```

Read and clean snp dataset
```{r message = FALSE}
snp_df =
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  # break up the variable date into integer variable year, month, and day
  separate(date, c("month", "day", "year"), convert = TRUE) %>%
  # replace month number with month name
  mutate (month = month.name[month]) %>%
  # remove day variable
  select(-day) %>%
  # organize year and month as the leading columns
  relocate(year, month)

# preview snp_df
snp_df
```

Read and clean the unemployment datase
```{r message = FALSE}
unemployment_df =
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>%
  # switch from “wide” to “long” format
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment_rate"
  ) %>%
  # convert abbreviated to full month name
  mutate(month = match(month, month.abb)) %>%
  mutate(month = month.name[month]) %>%
  janitor::clean_names() %>%
  # convert year to character variable
  mutate(year = as.integer(year))

# preview unemployment_df
unemployment_df
```

Merge the above three datasets
```{r}
# merge snp_df into pols_month_df
merge_pols_snp = left_join(pols_month_df, snp_df, by = c("year", "month"))
# merge unemployment_df into merge_pols_snp
merge_df = 
  left_join(merge_pols_snp, unemployment_df, by = c("year", "month")) %>%
  relocate(year, month, president, close, unemployment_rate)

# preview merge_df
merge_df
```

For pols_month_df, the dimension is `r nrow(pols_month_df)` rows * `r ncol(pols_month_df)` columns. The year ranges from `r min(pols_month_df$year)` to `r max(pols_month_df$year)`. The variables include: `r colnames(nyctransit_df)`. The president variable informs whether the president was democratic or republican. The values of the following variables inform the number of republican/democratic governors/senators/representatives on the associated date.

For snp_df, the dimension is `r nrow(snp_df)` rows * `r ncol(snp_df)` columns. The year ranges from `r min(snp_df$year)` to `r max(snp_df$year)`. The close variable informs the closing values of the S&P stock index on the associated date.

For unemployment_df, the dimension is `r nrow(unemployment_df)` rows * `r ncol(unemployment_df)` columns. The year ranges from `r min(unemployment_df$year)` to `r max(unemployment_df$year)`. The unemployment rate informs the percentage of unemployment in certain month of the associated year.

For the merge_df, the dimension is `r nrow(merge_df)` rows * `r ncol(merge_df)` columns. The year ranges from `r min(merge_df$year)` to `r max(merge_df$year)`. It is a merged dataset which informs the number of national politicians, the closing value, and unemployment rate. 