---
title: "p8105_hw2_wc2750"
author: "Weixi Chen"
date: "9/27/2020"
output: github_document
---

```{r message = FALSE}
# load library
library(tidyverse)
library(readxl)
library(dplyr)
```

## Problem 1

First, define a path to the dataset.
```{r}
path_to_data = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx"
```

Read the Mr. Trashwheel dataset.
```{r}
trashwheel_df = 
  read_xlsx(path = path_to_data,
            sheet = "Mr. Trash Wheel",
            range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data for 2018 and 2017
```{r}
precip_2018 = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet = "2018 Precipitation",
             skip = 1
            ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate (year = 2018) %>%
  relocate (year)

precip_2017 = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet = "2017 Precipitation",
             skip = 1
            ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate (year = 2017) %>%
  relocate (year)
```

Now combine annual precipitation
```{r}
month_df =
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017)

left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. The total precipitaion in 2018 is `r sum(precip_2018$total)`. The median number of sports balls in a dumpster in 2017 is `r median(trashwheel_df$sports_balls[trashwheel_df$year == 2017])`.

## Problem 2

Read and clean the NYC Transit dataset.
```{r message = FALSE}
nyctransit_df = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, route1:route11, entrance_type, entry, vending, ada) %>%
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

The NYC Transit dataset includes variables: `r colnames(nyctransit_df)`. As so far, I've cleaned the columns' names, selected the wanted columns, and converted the entry variable from character (YES vs NO) to a logical variable. The dimension of this dataset is `r nrow(nyctransit_df)` rows * `r ncol(nyctransit_df)` columns.

Additional information about NYC Transit dataset.
```{r}
# reformat distinct stations dataset
diststation_df = distinct(nyctransit_df, line, station_name, .keep_all = TRUE)
# number of distinct stations
nrow(diststation_df)
# number of stations are ADA compliant
nrow(filter(diststation_df, ada == "TRUE"))
# proportion of station entrances/exits without vending allow entrance
sum(nyctransit_df$vending == "NO" & nyctransit_df$entry == "TRUE")/sum(nyctransit_df$vending == "NO")
```

Reformat data with distinct route number and route name
```{r}
# reformat distinct route number and route name dataset for distinct stations
route_diststation_df = diststation_df %>%
  # convert route8:route11 to character variable
  mutate_at(vars(route8:route11), as.character) %>% 
  pivot_longer(
  route1:route11,
  names_to = "route",
  values_to = "train")
```

Additional information about serving A train
```{r}
# number of distinct stations serve the A train
nrow(filter(route_diststation_df, train == "A"))
# number of distinct stations serve the A trian and are ADA compliant
nrow(filter(route_diststation_df, train == "A" & ada == "TRUE"))
```

